A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-03-24 controlled_study.tex
box around the implications
back to 8+1 pages

Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it is bundled with every WordPress installation. According to the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively. These properties satisfy R1, our requirements for using a representative API. By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API. The API implements a self-documenting feature where API developers expose API endpoints over HTTP OPTIONS verb to explain the API elements. To implement this feature, the API developers describe the API elements in the code. For example:

@@

\begin{figure}[h!]
\begin{center}
\fbox{\includegraphics[width=\linewidth,keepaspectratio]{schema.png}}
\end{center}
\caption{Screenshot showing auto-generated documentation for Post/date}
\label{fig:schema}
\end{figure}

@@

\begin{itemize}
  \item Currently working as a software engineer.
  \item At least 1 year of industry experience as a software engineer.
  \item At least 1 year of industry experience with REST APIs.
  \item No prior experience with WordPress REST API.
\end{itemize}

@@

\textbf{Implication 1.} REST API documentation needs to include examples of data types such as Integer, String, and Array for each API field to satisfy API client developer information need.

@@

\textbf{Implication 2.} REST API developers need to include examples showing the valid data format for the API elements.

@@

\textbf{Implication 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.

@@

\textbf{Implication 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. Practitioners can follow our implications as a set of guidelines while documenting REST APIs to improve API usability.

>>>>

Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it is bundled with WordPress installation. According to the code repository, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively. These properties satisfy R1, our requirements for using a representative API. By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API. The API implements a self-documenting feature where API developers expose API endpoints over HTTP OPTIONS verb to explain the API elements. To implement this feature, the API developers describe the API elements in the code. For example:

@@

\begin{figure}[h!]
\begin{center}
\fbox{\includegraphics[width=\linewidth,keepaspectratio]{schema.png}}
\end{center}
\caption{Screenshot of documentation for Post/date}
\label{fig:schema}
\end{figure}

@@

\begin{itemize}
  \item Currently working as a software engineer.
  \item At least 1 year of industry experience as a developer.
  \item At least 1 year of industry experience with REST APIs.
  \item No prior experience with WordPress REST API.
\end{itemize}

@@

\hspace{-5mm}
\fbox{
\begin{minipage}{0.95\linewidth}

\textbf{Recommendation 1.} REST API documentation needs to include examples of data types such as Integer, String, and Array for each API field to satisfy API client developer information need.

\end{minipage}
}
\vspace{0.5mm}

@@

\vspace{1mm}
\hspace{-5mm}
\fbox{
\begin{minipage}{0.95\linewidth}

\textbf{Recommendation 2.} REST API developers need to include examples showing the valid data format for the API elements.
\end{minipage}
}
\vspace{0.5mm}

@@

\vspace{1mm}
\hspace{-5mm}
\fbox{
\begin{minipage}{0.95\linewidth}

\textbf{Recommendation 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
\end{minipage}
}
\vspace{0.5mm}

@@

\vspace{1mm}
\hspace{-5mm}
\fbox{
\begin{minipage}{0.95\linewidth}

\textbf{Recommendation 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
\end{minipage}
}
\vspace{0.5mm}

@@

\section{Conclusion}
In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples. We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body. We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API. Practitioners can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability.
